---
title: "Spatial data pipeline"
output: html_notebook
---

### Imports
```{r message=FALSE, warning=TRUE}
library(lubridate)
library(cbsodataR)
library(httr)
library(aweek)
library(glue)
library(tidyverse)
library(openxlsx)
library(sf)
library(fuzzyjoin)
```

### Get Population data from CBS

```{r}
pop.cbs <- cbs_get_data("37230NED") %>%
  filter(
    str_detect(Perioden, "\\d{4}MM\\d{2}") & 
    str_detect(RegioS, "(GM\\d+)|(NL01)")
  )

pop.all <- pop.cbs %>%
  mutate(
    Year = Perioden %>% str_extract("\\d{4}") %>% as.numeric(),
    Month = Perioden %>% str_match("\\d{4}MM(\\d{2})") %>% .[,2] %>% as.numeric(),
    Deaths = Overledenen_3,
    Population = BevolkingAanHetEindeVanDePeriode_15,
    Date = ISOdate(Year, Month, 1),
    Date = date(Date + duration(days = days_in_month(Date) - 1)),
    RegioS = ifelse(str_detect(RegioS, "NL01"), "NL", RegioS)
  ) %>%
  select(Date, "Year", "Month", "RegioS", "Population", "Deaths") %>%
  drop_na(Population) %>%
  arrange(desc(Year), desc(Month))

pop <- pop.all %>%
  mutate(Week = ISOweek::ISOweek(Date)) %>%
  group_by(RegioS, Week) %>%
  summarise(
    Population = last(Population)
  )

head(pop)
```
### Mortality

#### Get NL mortality data from CBS

```{r}

cbs_to_isoweek <- function(x) x %>%
  str_match("(\\d{4})[^J][0,1](\\d+)") %>% as.data.frame() %>% 
  set_names(c("Input", "Year", "Week")) %>%
  mutate(
    Week = Week %>% as.numeric(),
    Year = Year %>% as.numeric(),
    YearWeek = case_when(
      is.na(Week) ~ as.Date(NA),
      Week == 0 ~ date(ISOdate(Year, 1, 1)),
      Week == 1 ~ aweek::get_date(year = Year, week = 1),
      TRUE ~ aweek::get_date(year = Year, week = pmax(1, Week))
    ) %>% ISOweek::ISOweek() 
  ) %>% pull(YearWeek)

mort.nl <- cbs_get_data("70895NED") %>%
  filter(
    Geslacht == "1100" &
    LeeftijdOp31December == "10000"
  ) %>%
  mutate(
    YearWeek = cbs_to_isoweek(Perioden),
  ) %>%
  drop_na() %>%
  group_by(YearWeek) %>%
  summarise(
    Deaths = sum(Overledenen_1),
    Perioden = last(Perioden)
  ) %>% mutate(
    Year = str_match(YearWeek, "(\\d{4})") %>% .[,2] %>% as.numeric(),
    Week = str_match(YearWeek, "W(\\d+)") %>% .[,2],
    Date = aweek::get_date(year = Year, week = Week, day = 7),
    RegioS = "NL"
  ) %>% filter(Year >= 1995) %>% 
  select(Date, RegioS, Deaths, Week = YearWeek)

mort.nl
```

#### Get regional weekly mortality for 2020 and 2019

```{r}

areas <- GET("https://opendata.arcgis.com/datasets/e1f0dd70abcb4fceabbc43412e43ad4b_0.geojson") %>% content(as = "text") %>%
  read_sf() %>%
  rename(RegioS = Gemeentecode) %>% 
  arrange(Gemeentenaam) 

cur_week <- isoweek(today() - 7)

is_equal_or_startswith = ~ .x == .y | str_starts(.x, paste0(.y, "\\s"))

mort.weekly <- read.xlsx("https://www.cbs.nl/-/media/_excel/2020/{cur_week}/asd-tabellenset-sterfte-per-week-week24.xlsx" %>% glue(), 
          sheet = 3, detectDates = T, startRow = 4, colNames = F) %>% t() %>% as.tibble() %>%
  fill(`1`, .direction = "down") %>% select(-`3`) %>% set_names(c("Year", "Week", .[1,-(1:2)])) %>% .[-1,] %>%
  gather("Gemeentenaam", "Deaths", -Year, -Week) %>%
  group_by(Year, Gemeentenaam) %>%
  mutate(
    RowN = row_number(),
    WeekN = paste0("W", str_pad(RowN, 2, pad = "0")),
    Week = paste(Year, WeekN, sep = "-"),
    Deaths = Deaths %>% as.numeric(),
    Date = aweek::get_date(week = RowN, year = Year, day = 7)
  ) %>%
  ungroup() %>% drop_na(Deaths) %>%
  fuzzy_left_join(areas, c("Gemeentenaam" = "Gemeentenaam"), match_fun = is_equal_or_startswith) %>% select(Date, Week, RegioS, Deaths)

mort.weekly %>% filter(RegioS == "GM0345")
```

#### Interpolate monthly data pre 2019

Our data interval is weekly, so we will (perhaps not completely ethic) interpolated monthly death rate to weeks.

```{r}

days <- data.frame(Date = seq(min(pop.all$Date), to = max(pop.all$Date), by = "days")) %>%
  mutate(Week = ISOweek::ISOweek(Date)) %>% 
  crossing(pop.all %>% filter(RegioS != "NL") %>% distinct(RegioS))

mort.interpolated <- pop.all %>% arrange(Date) %>% 
  filter(RegioS != "NL") %>%
  group_by(RegioS) %>%
  mutate(
    DeathsAccum = cumsum(Deaths),
  ) %>% ungroup() %>% 
  full_join(days, c("RegioS", "Date")) %>% 
  arrange(Date) %>%
  group_by(RegioS) %>%
  mutate(
    DeathsAccumInt = approx(Date, DeathsAccum, xout=Date)$y,
    # DeathsAccumInt = replace(na.spline(DeathsAccum), is.na(na.approx(DeathsAccum, na.rm = F)), NA),
    Deaths2 = DeathsAccumInt - lag(DeathsAccumInt)
  ) %>%
  group_by(RegioS, Week) %>% summarise(
    Date = max(Date),
    Deaths = sum(Deaths2)
  ) %>%
  mutate(
    Deaths = if_else(Deaths < 0, 0, Deaths),
  )

ggplot(mort.interpolated %>% filter(RegioS == "GM0005"), aes(Date, Deaths)) +
  geom_line()
```

#### Bring them together

```{r}
mort <- mort.interpolated %>% full_join(mort.weekly, c("RegioS", "Week")) %>%
  mutate(Deaths = if_else(is.na(Deaths.y), Deaths.x, Deaths.y), Date = pmax(Date.x, Date.y, na.rm = T)) %>%
  union_all(mort.nl) %>%
  select("RegioS", "Week", "Date", "Deaths")

mort %>% filter(RegioS == "GM0345") %>% arrange(desc(Week))
```
### Age

```{r}
regios <- c("NL01  ", areas %>% st_drop_geometry() %>% distinct(RegioS) %>% pull(RegioS)) %>% 
  split(cut(seq_along(.), 20, labels = F))

ages <- c(seq(17000, 18400, 100), 22000) %>% as.character() %>% split(cut(seq_along(.), 3, labels = F))

pop.elder.cbs <- cross2(regios, ages) %>% 
  map(~cbs_get_data("03759ned", Geslacht = "T001038", BurgerlijkeStaat = "T001019", Leeftijd = .x[[2]], RegioS = .x[[1]])) %>%
  bind_rows() 

pop.elder <- pop.elder.cbs %>% filter(str_detect(Perioden, "\\d{4}JJ\\d{2}")) %>%
  group_by(RegioS, Perioden) %>%
  summarise(
    PopElder = sum(BevolkingOp1Januari_1)
  ) %>%
  ungroup() %>%
  mutate(
    Year = str_match(Perioden, "\\d{4}") %>% as.integer(),
    Date = date(ISOdate(Year, 1, 1)),
    Week = ISOweek::ISOweek(Date),
    RegioS = if_else(RegioS == "NL01  ", "NL", RegioS)
  ) %>% select(Week, RegioS, PopElder)

pop.elder %>% filter(RegioS == "NL")
```

### Get environment data

```{r}
temp <- read_rds("../results/temperature.Rds")
air <- read_rds("../results/air.Rds")

pop
```

### Join all into one dataframe

All tables are being join on year and month/week. 
Empty population data is filled from the previous months.

```{r}
data.joined <- mort %>%
  full_join(pop, by = c("RegioS", "Week")) %>%
  left_join(pop.elder, by = c("RegioS", "Week")) %>%
  left_join(temp, by = c("Week", "RegioS")) %>%
  left_join(air, by = c("Week", "RegioS")) %>%
  group_by(RegioS) %>%
  fill(Population, .direction = "updown")  %>%
  fill(PopElder, .direction = "updown") %>%
  ungroup() %>%
  group_by(Week) %>%
  mutate(
    Days = (max(Date) - min(Date) + 1) %>% as.numeric(units = "days"),
  ) %>%
  ungroup() %>%
  arrange(desc(Week))

head(data.joined, 5)
```

### Calculate the mortality rate

We will calculate the mortality rate the deaths by 1,000,000 inhabitants.
Additionally, we'll add columns for mortality rates in the previous 8 weeks.


```{r}

data <- data.joined %>%
  mutate(MortalityRate = Deaths * 100000 / Population, # to normalise week 53 and 1 
         PopElder = PopElder / Population,
         Interval = "week",
         Period = str_match(Week, "W(\\d+)") %>% .[,2] %>% as.numeric(),
         Year = str_match(Week, "(\\d{4})") %>% .[,2] %>% as.numeric()
  ) %>%
  group_by(RegioS) %>%
  mutate(
         "MortalityRate.-1" = lag(MortalityRate, 1, order_by = Date),
         "MortalityRate.-2" = lag(MortalityRate, 2, order_by = Date),
         "MortalityRate.-3" = lag(MortalityRate, 3, order_by = Date),
         "MortalityRate.-4" = lag(MortalityRate, 4, order_by = Date),
         "MortalityRate.-5" = lag(MortalityRate, 5, order_by = Date),
         "MortalityRate.-6" = lag(MortalityRate, 6, order_by = Date),
         "MortalityRate.-7" = lag(MortalityRate, 7, order_by = Date),
         "MortalityRate.-8" = lag(MortalityRate, 8, order_by = Date),
         "MortalityRate.-9" = lag(MortalityRate, 5, order_by = Date),
         "MortalityRate.-10" = lag(MortalityRate, 6, order_by = Date),
         "MortalityRate.-11" = lag(MortalityRate, 7, order_by = Date),
         "MortalityRate.-12" = lag(MortalityRate, 8, order_by = Date),
        ) %>%
  ungroup() %>%
  select(-Days)
data %>% filter(RegioS == "GM0345") %>% arrange(Week)
```

### Save data

```{r}
write_rds(data, "../results/data.Rds")
```


