---
title: "Air quality by area"
output: html_notebook
---

```{r message=FALSE, warning=FALSE}
options("rgdal_show_exportToProj4_warnings"="none")
library(rgdal)
library(tmaptools)
library(sf)
library(sp)
library(tmap)
library(httr)
library(lubridate)
library(tidyverse)
library(stringr)
require(raster)
```


### Get gemeente boundaries

```{r echo=FALSE, message=FALSE, warning=FALSE}
areas <- GET("https://opendata.arcgis.com/datasets/e1f0dd70abcb4fceabbc43412e43ad4b_0.geojson") %>% content(as = "text") %>%
  read_sf() %>%
  rename(RegioS = Gemeentecode) %>%
  select("RegioS")
```

### Get stations with coords.

```{r}
source("../../R/luchtmeetnet.R")

air.stations <- luchtmeetnet_get_data("stations", params = list()) %>%
  mutate(
    Coords = number %>% map(function(x) luchtmeetnet_get_data("stations/{x}" %>% glue(), params = list())$geometry$coordinates),
    Lat = Coords %>% map(first) %>% as.numeric(),
    Lon = Coords %>% map(last) %>% as.numeric(),
    Station = number
  ) %>% 
  st_as_sf(coords = c("Lat", "Lon")) %>%
  st_set_crs(4326) %>%
  st_join(areas)
```

#### Plot with locations

```{r purl=F}
tm_shape(areas) +
  tm_borders() +
  tm_shape(air.stations) +
  tm_dots("red")
```

### Add missing data 

```{r}
air.historic <- read_rds("../../results/luchtmeetnet.Rds")

# with no previous data this may take very very long. 
need_more <- ymd(max(air.historic$Date) + 1) <  today()

if (need_more) {
  t1 = now()
  dates <- seq(ymd(max(air.historic$Date) + 1), to = today(), by = "7 days")
  
  air.luchtmeetnet <- dates %>% map(function(startx)
    luchtmeetnet_get_data("measurements", params = list(
      start = format(startx, "%Y-%m-%dT%H:%M:%S"),
      end = format(startx + duration(days = 6, hours = 23, minutes = 59), "%Y-%m-%dT%H:%M:%S")
    ), verbose = TRUE)
  ) %>% bind_rows()  
  
  t2 = now()
  
  air.daily <- air.luchtmeetnet %>% 
    mutate(
      Date = date(ymd_hms(timestamp_measured)),
      Variable = if_else(formula == "PM25", "PM2.5", formula),
      Station = station_number
    ) %>%
    group_by(Date, Station, Variable) %>%
    summarise(
      Value = mean(value)
    ) %>% ungroup() %>%
    union_all(air.historic) %>% distinct() %>% arrange(Date)
  
  write_rds(air.daily, "../../results/luchtmeetnet.Rds")
} else {
  air.daily <- air.historic
}

head(air.daily)
```

#### Sample table

```{r paged.print=TRUE}
air.daily
```

### Aggregate by Week

```{r}
air.weekly <- air.daily %>%
  left_join(air.stations) %>%
  mutate(
    Year = year(Date),
    Month = month(Date),
    Week = ISOweek::ISOweek(Date),
  ) %>%
  group_by(Week, Station, Variable) %>%
  summarise(
    Value = mean(Value),
  ) %>%
  ungroup() %>%
  filter(Variable %in% c("PM10", "PM2.5", "NO2", "O3"))

air.weekly
```

### Transform all to planar and with km as units

```{r}
crs.grid <- "+proj=aea +lat_0=0 +lon_0=0 +lat_1=29.5 +lat_2=42.5 +x_0=0 +y_0=0 +ellps=GRS80 +units=km +no_defs +datum=WGS84"

areas.planar <- areas %>% st_transform(crs.grid)
air.planar <- air.stations %>% left_join(air.weekly)  %>% st_transform(crs.grid)
```

### Create grid

```{r}
grid <- raster::raster()
raster::extent(grid) <- raster::extent(areas.planar)
raster::res(grid) <- 1 # 5 km
raster::projection(grid) <- raster::crs(areas.planar)
```

### Interpolate

Different methods have been compared. The best is probably ordinary Kriging. Nevertheless, Inverted Distance Weights resulted a bit more faster and simpler approach.

```{r echo=FALSE, warning=FALSE}
interpolate_idw <- function(locations) {
  idm <- gstat::gstat(formula = Value ~ 1, locations = locations %>% st_as_sf())
  capture.output(idp <- raster::interpolate(grid, idm) %>% raster::mask(areas.planar))

  values <- tabularaster::cellnumbers(idp, areas.planar) %>% 
    mutate(v = raster::extract(idp, cell_)) %>% 
    group_by(object_) %>% summarise(Value = mean(v)) %>% pull(Value)
  
  return (areas.planar %>% mutate(
    Value = values,
    Week = first(locations$Week),
    Variable = first(locations$Variable)
  ))
}

plan(multiprocess)

air.interpolated <- air.planar %>% drop_na(Week, Value) %>% 
  select(-location, -number, -Coords, -Station, -RegioS) %>% as.data.frame() %>% 
  group_by(Week, Variable) %>% group_split() %>%
  future_map(interpolate_idw, .options = future_options(packages = c("sf"))) %>%
  bind_rows() %>% as.data.frame() %>% select(-geometry)

head(air.interpolated)
```

#### Example result

```{r purl=F}
tm_shape(air.interpolated %>% filter(Week == "2017-W10") %>% left_join(areas) %>% st_as_sf()) +
  tm_borders(col = "#fff1d5") +
  tm_fill(col = "Value") +
  tm_facets(by = "Variable")
```

### Add NL

```{r}
air.nl <- air.daily %>%
  mutate(
    Week = ISOweek::ISOweek(Date),
  ) %>%
  group_by(Week, Variable) %>%
  summarise(
    Value = mean(Value),
  ) %>%
  ungroup() %>%
  filter(Variable %in% c("PM10", "PM2.5", "NO2", "O3")) %>%
  mutate(
    RegioS = "NL"
  )

head(air.nl)
```

```{r}
air <- union_all(air.nl, air.interpolated) %>% spread(Variable, Value)
```
#### Write results

```{r}
write_rds(air, "../../results/air.Rds")
```

